% !TeX root = ssd.tex

\renewenvironment{abstract}{\small \list{}{%
    \setlength{\leftmargin}{2em}%
    \setlength{\rightmargin}{\leftmargin}%
    \listparindent 1.4em%
  }\item\relax{\bf Abstract.}}{\endlist}
  
\begin{abstract}
In this paper we study the two player randomized communication
complexity of the sparse set disjointness and the exists-equal 
problems and give matching lower and upper bounds for both. 
In the sparse set disjointness problem, each player
receives a $k$-subset of $[m]$ and the goal is to determine
whether the sets intersect. For this problem, we give a 
protocol that communicates a total of $O(k\log^{(r)}k)$ bits 
over $r$ rounds and errs with very small probability.
Taking $r=\log^{*}k$ one obtains a $O(k)$ total
communication $\log^{*}k$-round protocol with exponentially
small error probability, improving on the 
$O(k)$-bits $O(\log k)$-round constant error probability 
protocol of Håstad and Wigderson from 1997.

In the exists-equal problem, the players receive vectors 
$x,y\in[t]^n$ with the goal of determining whether there exists
a coordinate $i$ such that $x_i=y_i$ (i.e., OR of $n$ 
equality problems).
Exists-equal is an instance of sparse set disjointness with
$k=n$, hence the protocol above applies here as well, giving 
an $O(n\log^{(r)}n)$ upper bound.
Our main technical contribution is a matching 
lower bound: when $t=\Omega(n)$, any $r$-round
randomized protocol for the exists-equal problem with error 
probability at most $1/3$ should have a message of size 
$\Omega(n\log^{(r)}n)$. Our lower bound holds even for 
super-constant $r\le \log^*n$, showing that any $O(n)$ bits 
exists-equal protocol should have $\log^*n - O(1)$ rounds. 

The protocol we give errs only with less than 
polynomially small probability and provides guarantees on the 
total communication for the harder set disjointness problem, 
whereas our lower bound holds even for constant error 
probability protocols and for the easier exists-equal problem 
with guarantees on the maximum communication. Hence our upper and 
lower bounds match in a strong sense.
Our lower bound on the constant round protocols for
exists-equal shows that solving the OR of $n$ instances of the 
equality problems requires strictly more than $n$ times the cost 
of a single instance. To our knowledge this is the first example 
of such a {\em super-linear} increase in complexity.
\end{abstract}
